# Student Outcomes {#sec-so}

```{r setup, include=FALSE}
library(tidyverse)
library(tinytable)
library(readxl)
get_file <- function(url, outcome, indicator, course) {
  folder <- file.path("assessments", outcome, indicator) 
  dir.create(folder, recursive = TRUE, showWarnings = FALSE)
  download.file(url, file.path(folder, paste0(course, ".pdf")))
}

read_assessment_data <- function(file, ...) {
  normalize_names <- function(x) {
    x |>
      str_replace_all("\u00a0", " ") |>
      str_squish() |>
      str_to_lower()
  }

  standardize_names <- function(x) {
    normalized <- normalize_names(x)
    case_when(
      normalized == "id" ~ "id",
      normalized == "start time" ~ "start_time",
      normalized == "completion time" ~ "completion_time",
      normalized == "email" ~ "email",
      normalized == "name" ~ "name",
      normalized == "course" ~ "course",
      normalized == "semester" ~ "semester",
      normalized == "assessment" ~ "assessment",
      normalized == "indicator" ~ "indicator",
      str_detect(normalized, "^which course is this assessment given in\\?") ~ "course",
      str_detect(normalized, "^provide a brief description") ~ "description",
      str_detect(normalized, "^upload a pdf document") ~ "prompt",
      str_detect(normalized, "^which performance indicator") ~ "indicator",
      TRUE ~ x
    )
  }

  data <- readr::read_csv(file, ..., show_col_types = FALSE)
  normalized_names <- normalize_names(names(data))
  score_columns <- tibble(
    column = names(data),
    performance = case_when(
      str_detect(normalized_names, "^exceeds") ~ "exceeds",
      str_detect(normalized_names, "^meets") ~ "meets",
      str_detect(normalized_names, "^developing") ~ "developing",
      str_detect(normalized_names, "^unsatisfactory") ~ "unsatisfactory",
      TRUE ~ NA_character_
    )
  ) |>
    filter(!is.na(performance))

  collapse_scores <- function(performance_level) {
    columns <- score_columns |>
      filter(performance == performance_level) |>
      pull(column)

    if (length(columns) == 0) {
      return(rep(NA_real_, nrow(data)))
    }

    data |>
      transmute(across(all_of(columns), ~ readr::parse_number(as.character(.x)))) |>
      reduce(coalesce)
  }

  metadata_columns <- c(
    "id", "start_time", "completion_time", "email", "name", "course",
    "semester", "description", "assessment", "prompt", "indicator"
  )

  data |>
    select(-all_of(score_columns$column)) |>
    rename_with(standardize_names) |>
    select(any_of(metadata_columns)) |>
    bind_cols(tibble(
      exceeds = collapse_scores("exceeds"),
      meets = collapse_scores("meets"),
      developing = collapse_scores("developing"),
      unsatisfactory = collapse_scores("unsatisfactory")
    )) |>
    filter(if_any(everything(), ~ !is.na(.x) & as.character(.x) != ""))
}

read_factor_assessment_data <- function(file, ...) {
  normalize_names <- function(x) {
    x |>
      str_replace_all("\u00a0", " ") |>
      str_squish() |>
      str_to_lower()
  }

  standardize_names <- function(x) {
    normalized <- normalize_names(x)
    case_when(
      normalized == "id" ~ "id",
      normalized == "start time" ~ "start_time",
      normalized == "completion time" ~ "completion_time",
      normalized == "email" ~ "email",
      normalized == "name" ~ "name",
      str_detect(normalized, "^which course is this assessment given in\\?") ~ "course",
      str_detect(normalized, "^provide a brief description") ~ "description",
      str_detect(normalized, "^upload a pdf document") ~ "prompt",
      str_detect(normalized, "^which performance indicator") ~ "indicator",
      str_detect(normalized, "^enter the total number of students") ~ "total",
      TRUE ~ x
    )
  }

  factor_labels <- c(
    "public health1" = "Public health",
    "safety1" = "Safety",
    "welfare" = "Welfare",
    "global issues" = "Global",
    "cultural issues" = "Cultural",
    "social issues" = "Social",
    "environmental issues" = "Environmental",
    "economic issues" = "Economic"
  )

  data <- readr::read_csv(file, ..., show_col_types = FALSE)
  normalized_names <- normalize_names(names(data))
  factor_columns <- tibble(
    column = names(data),
    normalized = normalized_names
  ) |>
    filter(normalized %in% names(factor_labels)) |>
    mutate(design_factor = factor_labels[normalized])

  if (nrow(factor_columns) == 0) {
    stop("No design factor assessment columns were found in ", file, call. = FALSE)
  }

  metadata_columns <- c(
    "id", "start_time", "completion_time", "email", "name", "course",
    "description", "prompt", "indicator", "total"
  )

  data |>
    rename_with(standardize_names) |>
    filter(str_detect(indicator, regex("Consider factors", ignore_case = TRUE))) |>
    select(any_of(metadata_columns), all_of(factor_columns$column)) |>
    pivot_longer(
      cols = all_of(factor_columns$column),
      names_to = "factor_column",
      values_to = "meets"
    ) |>
    mutate(
      design_factor = factor_labels[normalize_names(factor_column)],
      total = readr::parse_number(as.character(total)),
      meets = readr::parse_number(as.character(meets)),
      unsatisfactory = total - meets
    ) |>
    select(any_of(metadata_columns), design_factor, meets, unsatisfactory) |>
    filter(!is.na(total), !is.na(meets))
}

format_table_for_output <- function(data) {
  if (!knitr::is_latex_output()) {
    return(data)
  }

  data |>
    mutate(across(where(is.character), ~ str_replace_all(.x, fixed("#"), "No. ")))
}

library(wesanderson)
pal <- wes_palette("AsteroidCity1", 4)[c(3, 2, 1, 4)]
theme_set(theme_minimal())
```

The program has seven student outcomes, which describe abilities and skills
that students should demonstrate upon graduation. 

1. An ability to identify, formulate, and solve complex engineering problems by applying principles of engineering, science, and mathematics.
2. An ability to apply engineering design to produce solutions that meet specified needs with consideration of public health, safety, and welfare as well as global, cultural, social, environmental, and economic factors.
3. An ability to communicate effectively with a range of audiences.
4. An ability to recognize ethical and professional responsibilities in engineering situations and make informed judgments, which must consider the impact of engineering solutions in global, economic, environmental, and societal contexts.
5. An ability to function effectively on a team whose members together provide leadership, create a collaborative and inclusive environment, establish goals, plan tasks, and meet objectives.
6. An ability to develop and conduct appropriate experimentation, analyze and interpret data, and use engineering judgment to draw conclusions.
7. An ability to acquire and apply new knowledge as needed, using appropriate learning strategies.

The primary assessment of these student outcomes is a direct assessment of
individual student work undertaken at multiple points in the program curriculum.
We attempt to formally assess each student outcome at two levels:

  1. At a Formative level, in lower-division courses. There may be more than one formative assessment of a student outcome.
  3. At a Summative level, in an upper-division course as close to graduation as possible.

Summative assessments are given priority in program evaluation, but formative
assessments can provide additional supportive data.  @tbl-socourses lists which
courses are used for formative and summative assessments. 
Formative assessments may not assess all indicators of a student
outcome, but summative assessments cover all indicators. Additionally, formative
assessments may include data collected from students in other programs.

```{=latex}
\let\quartoRegularTable\table
\let\endquartoRegularTable\endtable
\let\table\sidewaystable
\let\endtable\endsidewaystable
```

```{r}
#| label: tbl-socourses
#| tbl-cap: Student Outcome Assessment Locations
socourses <- tibble::tribble(
  ~Outcome, ~CCE102, ~CCE103, ~Software, ~Measurements, ~CCE203, ~CE232, ~CCE270, ~CE291, ~Breadth, ~Design, ~Capstone,
  "1. Complex problems",      "", "F", "", "", "F", "", "", "", "S",  "", "",
  "2. Design",                "", "", "", "", "", "", "", "", "", "F", "S",
  "3. Communication",         "", "", "F", "", "", "", "F", "", "", "", "S",
  "4. Ethics",                "F", "", "", "", "", "", "", "F", "", "", "S",
  "5. Teamwork",              "F", "", "", "", "", "", "", "F", "", "", "S",
  "6. Experiments and data",  "", "", "", "", "", "F", "F", "", "S", "", "",
  "7. Acquire knowledge",     "", "", "F", "", "", "", "", "F", "", "", "S"
)

socourses_table <- tt(socourses, 
  notes = list(a = list(i = 6, j = 10, text = "Only selected breadth courses contribute to the summative assessment."))
) |>
  style_tt(align = "l") |>
  theme_tt("default") 

socourses_table
```

```{=latex}
\let\table\quartoRegularTable
\let\endtable\endquartoRegularTable
```




We assess each outcome on a three-year schedule. @tbl-so-schedule shows two
cycles of the three-year assessment schedule. In a six-year ABET accreditation
cycle, we collect data on student performance for each outcome twice. Courses
are assessed in one of Fall or Winter semesters for courses that are offered in
both semesters. Early in the assessment semester, instructors are informed that
they will be expected to provide assessment data in their class. At the end of
the semester, instructors are invited to complete a form that requires a brief
description of the assessment (e.g., final exam question), a specific assessment
instrument as given to students, and the number of students assessed at each
level of the rubric. 

```{=latex}
\let\quartoRegularTable\table
\let\endquartoRegularTable\endtable
\let\table\sidewaystable
\let\endtable\endsidewaystable
```

```{r}
#| label: tbl-so-schedule
#| tbl-cap: Assessment Schedule

so_schedule <- tibble::tribble(
  ~Outcome, ~`25 Winter`, ~`25 Sp/Su`, ~`25 Fall`, ~`26 Winter`, ~`26 Sp/Su`, ~`26 Fall`, ~`27 Winter`, ~`27 Sp/Su`, ~`27 Fall`, ~`28 Winter`, ~`28 Sp/Su`,
  "SO1: Complex Problems", "", "", "", "", "", "A", "E", "", "C", "", "",
  "SO2: Design", "", "", "A", "A", "E", "C", "", "", "", "", "",
  "SO3: Communications", "", "", "", "", "", "A", "E", "", "C", "", "",
  "SO4: Ethics", "A", "E", "C", "", "", "", "", "", "", "A", "E",
  "SO5: Teamwork", "", "", "A", "A", "E", "C", "", "", "", "", "",
  "SO6: Experiments / Data", "A", "E", "C", "", "", "", "", "", "", "A", "E",
  "SO7: Acquire Knowledge", "", "", "A", "A", "E", "C", "", "", "", "", "",
  "ABET Schedule", "", "", "", "", "Self-study", "Visit", "Response", "", "", "", ""
)

so_schedule_table <- tt(so_schedule, notes = "A: Assessment, E: Evaluation, C: Change") |>
  style_tt(align = "l") 

so_schedule_table
```

```{=latex}
\let\table\quartoRegularTable
\let\endtable\endquartoRegularTable
```


For each of the seven student outcomes, the program develops a set of
performance indicators and an associated rubric that faculty apply to specific
and selected student assessments in the assessed courses. The levels of the
rubric are

  - Unsatisfactory 
  - Developing 
  - Meets Expectations
  - Exceeds Expectations
  
The performance indicators and rubric apply to all levels of instruction, though
the level of expected student performance increases as students grow and learn
through the program. 

For all indicators, the program goal is that 75 percent of students will
demonstrate performance that meets or exceeds expectations by graduation. The
following sections describe the performance indicators and associated rubrics
used to assess each student outcome, along with a brief description of the
assessment instruments used for each course and assessment and evaluation information. 
Specific assessment instruments are archived and available in the program assessment folder.

The annual program stewardship report includes the present academic year's
assessment and evaluation, as well as reports on actions taken related to the previous
year's recommended changes.



::: {.callout-note title="Assessment Practices" collapse="true"}
The program assessment coordinator creates a Microsoft Forms survey for each student outcome. In the semester
that a course is assessed, the program assessment coordinator sends the survey to the instructor at least twice:

  1. Before the semester begins, to remind the instructor of their assessment responsibilities and to invite them to consider
  activities that will be used for the assessment.
  2. At the end of the semester, to collect the final assessment data.

The survey asks instructors to provide **for each indicator**:
  
  - A brief description of the indicator; e.g., "Final exam question"
  - The assessment instrument as given to students; e.g. the text of the final exam question with accompanying diagrams or data. This must be provided as a PDF file.
  - The number of students assessed at each level of the rubric for the indicator.

The assessment coordinator collects the assessment data from the Forms databases and transfers them into this repository.
The coordinator cleans the data as necessary to create the assessment figures and commits the rendered page.
:::


::: {.callout-important title="Best Practices for Instructors" collapse="false"}
Activities used for student outcomes assessment should be clearly labeled on the
syllabus; i.e., "HW 4.1. This problem is used to assess student outcome X in our
program assessments." This accomplishes two things. First, it helps instructors
to understand what activities are used in assessment as the course is taught by
others. Second, it communicates to the students that their performance on the
activity is used for more than their own grade.

A single activity may be used to assess multiple student outcomes, or multiple
indicators of a single outcome. But each indicator should be assessed separately
based on the rubric for that indicator. For example,
a term project may provide assessment data on both "Communication: Write"
indicator and on the "Design: Recommend design" indicators. But student performance
on the indicator should be assessed separately, based on the rubric.

It is not required that the rubrics listed in this section be determinative of
the grade the students receive on the activity. That is, an instructor may still
elect to issue full marks for work assessed at a "Developing" level, as an
example. However, it is best practice to conform the rubric used in grading to
the rubric used in the assessment, so that students are fully informed of our
expectations for their performance. In the term project example above, an instructor
may choose to award more points for the "Communication: Write" indicator than
for the "Design: Recommend design" indicator based on their goals for the course.

Instructors should select work that students complete **independently** if at
all possible.  Team project reports or other collaborative efforts make it
difficult to assess student performance.  When assessing teamwork-related
indicators using team projects, instructors should make an effort to understand
how individual students performed.

Learning Suite contains tools that allow instructors to link assignments to course learning outcomes.
Doing this automates many features of collecting assessment data.
:::

## Complex Problems {#sec-so1}

Students will demonstrate:

> An ability to identify, formulate, and solve complex engineering problems by
> applying principles of engineering, science, and mathematics.​

:::{.callout-warning title="ABET Definition"}
The ABET definition of "Complex Problems" is 

> Complex engineering problems include one or more of the following
> characteristics: involving wide-ranging or conflicting technical issues, having
> no obvious solution, addressing problems not encompassed by current standards
> and codes, including many component parts or sub-problems, involving multiple
> disciplines, or having significant consequences in a range of contexts.

It is understood that just because a problem is "hard" does not mean it is complex. Similarly, a problem may have 
many sub-steps, but if the parts are not integrated or coordinated the problem would not be considered "complex."
:::


### Performance Indicators {#sec-so1-performance}

We evaluate this outcome based on three performance indicators. Students should:

1.	Identify the problem parameters and given values.
2.	Formulate a solution strategy.
3.	Evaluate the best alternative from many possible solutions.

We assess these performance indicators with the rubric in @tbl-so1-rubric.

| Indicator | Unsatisfactory | Developing | Meets Expectations | Exceeds Expectations |
|:---|:---|:---|:---|:---|
| 1. Identify | Fails to identify given parameters. | Identifies some given parameters. | Identifies all provided parameters. | Identifies all parameters including unreasonable parameters or those requiring assumptions. |
| 2. Formulate | Fails to formulate a solution strategy. | Formulates a strategy with limitations. | Formulates a solution strategy based in accepted methods. | Formulates a strategy that adjusts for the weaknesses of existing techniques. |
| 3. Evaluate | Fails to evaluate alternative solutions. | Evaluates an alternative solution. | Evaluates alternative solutions. | Evaluates alternatives with a consideration of human needs. |
: Student Outcome 1 Assessment Rubric {#tbl-so1-rubric}

### Assessment Instruments {#sec-so1-assessment}

We assess this student outcome in the following courses with the listed assessments:

Formative

-	CCE 103: Engineering Mechanics – Statics 
    -	Identify: Final exam question
    -	Formulate: Final exam question
-	CCE 203: Engineering Mechanics – Materials 
    -	Identify: Final exam question
    -	Formulate: Final exam question

Summative

-	CE 321: Structural Analysis 
    -	Identify: Truss analysis project
    -	Formulate: Truss analysis project
    -	Evaluate: Truss analysis project
-	CE 331: Hydrology 
    -	Identify: Term project
    -	Formulate: Term project
    -	Evaluate: Term project
-	CE 341: Soil Mechanics 
    -	Identify: Final exam question
    -	Formulate: Final exam question
    -	Evaluate: Final exam question
-	CE 351: Environmental Engineering 
    -	Identify: Final exam question
    -	Formulate: Final exam question
    -	Evaluate: Final exam question
-	CE 361: Transportation Engineering 
    -	Identify: Pavement management homework
    -	Formulate: Pavement management homework
    -	Evaluate: Pavement management homework


### Evaluation
We will assess and evaluate this student outcome in the 2026-27 academic year.

## Design and Contemporary Issues {#sec-so2}
Students will demonstrate:

> An ability to apply engineering design to produce solutions that meet
> specified needs with consideration of public health, safety, and welfare as well
> as global, cultural, social, environmental, and economic factors.


:::{.callout-warning title="ABET Definition"}
The ABET definition of "Design" is

 > Engineering design is a process of devising a system, component, or process
 > to meet desired needs and specifications within constraints. It is an
 > iterative, creative, decision-making process in which the basic sciences,
 > mathematics, and engineering sciences are applied to convert resources into
 > solutions. Engineering design involves identifying opportunities, developing
 > requirements, performing analysis and synthesis, generating multiple solutions,
 > evaluating solutions against requirements, considering risks, and making
 > trade-offs, for the purpose of obtaining a high-quality solution under the
 > given circumstances. For illustrative purposes only, examples of possible
 > constraints include aesthetics, codes, constructability, cost, ergonomics,
 > extensibility, functionality, interoperability, legal considerations,
 > maintainability, manufacturability, marketability, policy, regulations,
 > schedule, standards, sustainability, or usability.


:::

### Performance Indicators {#sec-so2-performance}
We evaluate this outcome based on four performance indicators. Students should:

1.	Develop the criteria against which the design will be judged. These criteria may include safety requirements, performance expectations, or industry standards.
2.	Select appropriate computational, experimental, or qualitative methods to develop the design.
3.	Justify the selection of a preferred solution from among multiple alternatives.
4.	Consider public health, safety, and welfare as well as global, cultural, social, environmental, and economic factors.

We assess these performance indicators with the rubric in @tbl-so2-rubric.

| Indicator | Unsatisfactory | Developing | Meets Expectations | Exceeds Expectations |
|:---|:---|:---|:---|:---|
| 1. Develop design criteria | Little recognition of criteria or selection of irrelevant criteria. | Partial recognition of relevant criteria. | Thorough recognition of relevant criteria. | Thorough recognition of relevant criteria, plus optional ones. |
| 2. Select methods | Poor matching of methods to problem. | Partial selection or justification of relevant methods. | Selection and justification of methods relevant to solving problem. | Selection and justification of relevant methods, acknowledging limitations. |
| 3. Recommend design | No design, wrong design, or unjustified design. | Partially acceptable design and justification. | Acceptable design justified by methods and criteria. | Acceptable, justified design chosen from multiple alternatives. |
| 4. Consider factors | Unsatisfactory |  | Meets Expectations |  |
| Public Health | Does not consider |  | Considers |  |
| Safety | Does not consider |  | Considers |  |
| Welfare | Does not consider |  | Considers |  |
| Global | Does not consider |  | Considers |  |
| Cultural | Does not consider |  | Considers |  |
| Social | Does not consider |  | Considers |  |
| Environmental | Does not consider |  | Considers |  |
| Economic | Does not consider |  | Considers |  |
: Student Outcome 2 Assessment Rubric {#tbl-so2-rubric}


### Assessment Instruments {#sec-so2-assessment}


```{r so2-clean, include = FALSE}
so2 <- read_assessment_data("data/Student Outcome 2_ Design(Sheet1).csv")
so2_factors <- read_factor_assessment_data("data/Student Outcome 2_ Design(Sheet1).csv") |>
  mutate(
      course = case_when(
      str_detect(course, "CE 421") ~ "CE 421",
      str_detect(course, "CE 424") ~ "CE 424",
      str_detect(course, "CE 472") ~ "CE 472",
      str_detect(course, "CE 453") ~ "CE 453",
      str_detect(course, "CE 461") ~ "CE 461",
      str_detect(course, "CE 433") ~ "CE 433",
    TRUE ~ NA_character_
    )
  )

so2_clean <- so2 |> 
  pivot_longer(exceeds:unsatisfactory, names_to = "Performance", values_to = "Number") |>
  filter(!is.na(Number)) |>
  mutate(
    Performance = factor(Performance, levels = c("exceeds", "meets", "developing", "unsatisfactory"), 
    labels = c("Exceeds Expectations", "Meets Expectations", "Developing", "Unsatisfactory")),
    indicator = case_when(
      str_detect(indicator, "Develop design criteria") ~ "Develop design criteria",
      str_detect(indicator, "Select methods") ~ "Select methods",
      str_detect(indicator, "Recommend design") ~ "Recommend design",
      str_detect(indicator, "Consider factors") ~ "Consider factors",
      TRUE ~ indicator
    ),
    course = case_when(
      str_detect(course, "CE 421") ~ "CE 421",
      str_detect(course, "CE 424") ~ "CE 424",
      str_detect(course, "CE 472") ~ "CE 472",
      str_detect(course, "CE 453") ~ "CE 453",
      str_detect(course, "CE 461") ~ "CE 461",
      str_detect(course, "CE 433") ~ "CE 433",
      TRUE ~ NA_character_
    ) 
  ) |> 
    group_by(course, indicator) |>
    mutate(total = sum(Number)) |>
    mutate(Percent = if_else(total > 0, Number / total * 100, NA_real_))

so2_factors_clean <- so2_factors |>
  mutate(
    design_factor = factor(
      design_factor,
      levels = c("Public health", "Safety", "Welfare", "Global", "Cultural", "Social", "Environmental", "Economic")
    )
  ) |>
  group_by(course, design_factor) |>
  summarise(
    meets = sum(meets, na.rm = TRUE),
    unsatisfactory = sum(unsatisfactory, na.rm = TRUE),
    total = sum(total, na.rm = TRUE),
    .groups = "drop"
  ) |>
  pivot_longer(c(unsatisfactory, meets), names_to = "Performance", values_to = "Number") |>
  mutate(
    Performance = factor(
      Performance,
      levels = c("unsatisfactory", "meets"),
      labels = c("Unsatisfactory", "Meets Expectations")
    ),
    Percent = if_else(total > 0, Number / total * 100, NA_real_)
  )
```

We assess this student outcome in the following courses with the assessments listed in @tbl-so2-instruments.
Complete instrument prompts are available in the program assessment folder.

```{r}
#| label: tbl-so2-instruments
#| tbl-cap: Assessment Instruments for Student Outcome 2
so2_instruments <- bind_rows(
  so2_clean |> 
  group_by(course, indicator, description) |>
  slice(1) |>
  mutate(level = ifelse(course == "CE 472", "Summative", "Formative"))  |>
  arrange(indicator, level, course) |>
  ungroup() |>
  select(indicator, level, course,  description)  ,
so2_factors |>
  select(course, description) |>
  group_by(course, description) |>
  slice(1) |>
  mutate(
    indicator = "Consider Factors", 
    level = ifelse(course == "CE 472", "Summative", "Formative")
  )  
) |>
  setNames(c("indicator", "Assessment Level", "Course", "Instrument")) |>
  format_table_for_output()

tt(so2_instruments)  |>
  group_tt(i = indicator) |>
  subset(select = -indicator)  |> 
  style_tt(align = "l") |>
  style_tt(
    i = c(1, 5, 12, 15),
    color = "white",
    background = "gray",
    bold = TRUE)

```



### Data and Evaluation

 @fig-s02 shows the share of students assessed at each performance level for each indicator in the relevant courses. The assessment is as follows:

  - **Develop design criteria**: Two formative assessments (CE 421 and CE 424) and one summative assessment (CE 472) 
  reveal that program students are meeting expectations at the 75% threshold level. 
  - **Select methods**: One formative assessment (CE 421) and one summative assessment (CE 472) 
  both assess 100% of students at the meeting expectations level.
  - **Recommend design**: Five formative assessments and one summative assessment (CE 472) vary in their assessment of
  program students. In CE 421 and 433, students are meeting expectations at the 75% threshold level. In the other courses --- especially in the summative assessment of CE 472 --- students are not meeting expectations at the 75% threshold.


```{r}
#| label: fig-s02
#| fig-cap: Fall 2025 and Winter 2026 Assessment of Student Outcome 2
#| fig-width: 7
ggplot(so2_clean, aes(x = Percent, y = course, fill = fct_rev(Performance))) + 
  geom_bar(stat = "identity") +
  labs(y = "Course", x = "Percent of students") +  
  facet_grid(~ indicator) + 
  geom_vline(xintercept = 75, linetype = "dashed", color = "red") +
  theme_bw() +
  theme(legend.position = "bottom") +
  scale_fill_manual("Performance", values = pal)
```

@fig-s02-factors shows the share of students who consider the various design factors in the "Consider factors" performance indicator. In the formative assessment of CE 421, all students were assessed as considering the factors appropriately. In CE 472, students met the 75% expectations threshold for Safety, Environmental, and Economic factors but not for the Public Health, Welfare, Global, Cultural, or Social factors.

```{r}
#| label: fig-s02-factors
#| fig-cap: Fall 2025 and Winter 2026 Assessment of Student Outcome 2 design factors.
#| fig-width: 7
ggplot(so2_factors_clean, aes(y = Percent, x = course, fill = Performance)) +
  geom_col() + 
  labs(x = "Course", y = "Percent of students") +
  facet_grid(~ design_factor) +
  geom_hline(yintercept = 75, linetype = "dashed", color = "red") +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(
    "Performance",
    values = c("Unsatisfactory" = pal[1], "Meets Expectations" = pal[3])
  )
```

The committee recommends that more design courses develop activities that will instruct students in topics related to "Develop design criteria," "Select methods,"  and "Consider factors" indicators and assess the activities. 


## Communication {#sec-so3}
Students will demonstrate:

> An ability to communicate effectively with a range of audiences.


### Performance Indicators 

We evaluate this outcome based on four performance indicators. Students should: 
1.	Write technical texts in clear and organized English 
2.	Apply industry conventions for professional reports (typography, reference format, etc.) 
3.	Use graphics to illustrate numeric data 
4.	Present information in oral contexts 

We assess these performance indicators with the rubric in @tbl-so3-rubric.

| Indicator | Unsatisfactory | Developing | Meets Expectations | Exceeds Expectations |
|:----------|:---------------|:-----------|:-------------------|:---------------------|
| 1. Write | Unclear, disorganized, or incorrect writing interferes with understanding. | Writing has some incorrect or unclear elements. | Writing is clear and well-organized. | Writing is exceptionally clear, organized, and coherent. |
| 2. Conventions | Does not apply industry conventions. | Applies some industry conventions. | Largely applies industry conventions. | Precisely applies industry conventions. |
| 3. Graphics | Graphics are missing or poorly constructed. | Graphics have some construction or clarity issues. | Graphics are clear and well-made. | Graphics are exceptionally well-produced and informative. |
| 4. Oral | Delivery is unclear or disorganized. | Delivery communicates basic ideas but lacks clarity or structure. | Delivery is clear, structured, and supports understanding. | Delivery is polished, concise, and highly effective. |
: Student Outcome 3 Assessment Rubric {#tbl-so3-rubric}



### Assessment Instruments


### Data and Evaluation



## Ethics {#sec-so4}
Students will demonstrate:

> An ability to recognize ethical and professional responsibilities in engineering situations and make informed judgments, which must consider the impact of engineering solutions in global, economic, environmental, and societal contexts.

### Performance Indicators {.unnumbered}

We evaluate this outcome based on three performance indicators. Students should:

1.	Recognize personal and professional responsibilities towards others including employees, clients, the wider profession, and the public.
2.	Recognize moral and ethical dilemmas in engineering situations and make judgments to identify appropriate resolutions.
3.	Consider the impacts of an engineering solution in global, economic, environmental, and societal contexts.

We assess these performance indicators with the rubric in @tbl-so4-rubric.

| Indicator                  | Unsatisfactory                                                                | Developing                                                                           | Meets Expectations                                                             | Exceeds expectations                                                                                                                |
|:---------------------------|:------------------------------------------------------------------------------|:-------------------------------------------------------------------------------------|:-------------------------------------------------------------------------------|:------------------------------------------------------------------------------------------------------------------------------------|
| 1. Recognize responsibilities | Does not recognize or identify professional responsibilities.                 | Recognizes some professional responsibilities, but misses at least one.              | Recognizes professional responsibilities towards all key stakeholders          | Demonstrates interest in and  motivation to serve society through the profession.                                                   |
| 2. Make judgments             | Does not identify ethical dilemmas, and or develops inappropriate resolutions | Misidentifies ethical dilemmas, and or develops inadequate resolutions.              |  Properly identifies ethical dilemmas and identifies appropriate resolutions.  | Actively considers ethical challenges that may arise in professional contexts and seeks to mitigate  them.                          |
| 3. Consider impacts           | Fails to identify impacts in any contexts.                                    | Identifies impacts in some contexts, but fails to identify likely impacts in others. | Identifies likely impacts in a variety  of contexts.                           | Actively seeks to consider the ethical impact of engineering situations in global, economic, environmental, and societal  contexts. |
: Student Outcome 4 Assessment Rubric {#tbl-so4-rubric}

### Assessment Instruments {.unnumbered}

We evaluated this outcome most recently in Winter 2025. All graduating students take CE 472 in Winter semester. Students take CE 291 R each semester from their sophomore year onward. CCE 102 may be taken in Fall or Winter semesters, so this is approximately a 50% sample of the students in the program. CCE 102 also includes students who are in the CFM and the FM programs.

```{r so4-clean, include = FALSE}
so4 <- read_assessment_data("data/Student Outcome 4_ Ethics, Faith and Sustainability(Sheet1).csv")
so4_clean <- so4 |> 
  pivot_longer(exceeds:unsatisfactory, names_to = "Performance", values_to = "Number") |>
  filter(!is.na(Number)) |>
  mutate(
    Performance = factor(Performance, levels = c("exceeds", "meets", "developing", "unsatisfactory"), 
    labels = c("Exceeds Expectations", "Meets Expectations", "Developing", "Unsatisfactory")),
    indicator = factor(indicator, labels = c("Consider Impacts", "Make Judgments", "Recognize Responsibilities")),
    course = case_when(
      str_detect(course, "CCE 102") ~ "CCE 102",
      str_detect(course, "CE 291") ~ "CE 291R",
      str_detect(course, "CE 472") ~ "CE 472",
      TRUE ~ NA_character_
    ) 
  ) |> 
    group_by(course, indicator) |>
    mutate(total = sum(Number)) |>
    mutate(Percent = if_else(total > 0, Number / total * 100, NA_real_))
```


We assess this student outcome in the following courses with the assessments listed in @tbl-so4-instruments.
Complete instrument prompts are available in the program assessment folder.

```{r}
#| label: tbl-so4-instruments
#| tbl-cap: Assessment Instruments for Student Outcome 2
so4_instruments <- so4_clean |> 
  group_by(course, indicator, description) |>
  slice(1) |>
  mutate(level = ifelse(course == "CE 472", "Summative", "Formative"))  |>
  arrange(indicator, level, course) |>
  ungroup() |>
  select(indicator, level, course,  description) |> 
  setNames(c("indicator", "Assessment Level", "Course", "Instrument")) |>
  format_table_for_output()

tt(so4_instruments)  |>
  group_tt(i = indicator) |>
  subset(select = -indicator)  |> 
  style_tt(align = "l") |>
  style_tt(
    i = c(1, 4, 8),
    color = "white",
    background = "gray",
    bold = TRUE)

```


### Data and Evaluation

We evaluated this outcome most recently in Winter 2025. All graduating students
take CE 472 in Winter semester. Students take CE 291 R each semester from their
sophomore year onward. CCE 102 may be taken in Fall or Winter semesters, so this
is approximately a 50% sample of the students in the program. CCE 102 also
includes students who are in the CFM and the FM programs.  @fig-s04 shows the
share of students at each performance level for each indicator for each course,
alongside the threshold of 75%. By taking CE 472, 75% of students meet the
expectations for each indicator.

```{r}
#| label: fig-s04
#| fig-cap: Winter 2025 Assessment of Student Outcome 4
#| fig-width: 7
ggplot(so4_clean, aes(y = Percent, x = course, fill = fct_rev(Performance))) + 
  geom_col() + 
  labs(x = "Course", y = "Percent of students") +
  facet_grid(~ indicator) + 
  geom_hline(yintercept = 75, linetype = "dashed", color = "red") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual("Performance", values = pal)
```

```{r get-files, eval = FALSE}
# download the files in the pdf document field
for(i in 1:nrow(so4)) {
  get_file(
    url = so4$prompt[i],
    outcome = "SO4",
    indicator = so4$indicator[i],
    course = so4$course[i]
  )
}

```

The evaluation of this assessment data led the instructors of CCE 102 and CE 472 to make changes in the
way that they taught and evaluated ethical dilemmas. The re-assessment of this
outcome will occur in the 2027-2028 academic year.


## Teamwork {#sec-so5}
Students will demonstrate:

> An ability to function effectively on a team whose members together provide leadership, create a collaborative and inclusive environment, establish goals, plan tasks, and meet objectives.

### Performance Indicators {#sec-so5-performance}

We evaluate this outcome based on three performance indicators. Students should: 

1. Show leadership through accountability and service.
2. Contribute to a collaborative and inclusive environment where all team members feel a sense of belonging.
3. Meet team objectives through establishing goals and planning tasks.
 
We assess these performance indicators with the rubric in @tbl-so5-rubric.

| Indicator | Unsatisfactory | Developing | Meets Expectations | Exceeds Expectations |
|:---|:---|:---|:---|:---|
| 1. Accountability and service | Does not accomplish assigned tasks. | Accomplishes assigned tasks but is delayed or requires reminders. | Accomplishes tasks on time. | Accomplishes tasks on time, and helps team members accomplish their roles. |
| 2. Contribute to team belonging | No or negative involvement in team activities. | Involved in team activities and makes some effort to engage others. | Actively involved in team activities and encourages engagement from others. | Drives team activities and motivates productive engagement from others. |
| 3. Meet team objectives | Fails to identify reasonable goals or appropriate tasks. | Develops goals and tasks that are infeasible, or that required substantial revision. | Develops goals and a task plan that results in meeting the assignment expectations. | Develops goals and a resilient task plan that results in exceeding the assignment expectations. |
: Rubric to assess student outcome 5: Teamwork {#tbl-so5-rubric}

### Assessment Instruments {#sec-so5-assessment}

```{r so5-clean, include = FALSE}
so5 <- read_assessment_data("data/Student Outcome 5_ Teamwork(Sheet1).csv")

so5_clean <- so5 |> 
  pivot_longer(exceeds:unsatisfactory, names_to = "Performance", values_to = "Number") |>
  filter(!is.na(Number)) |>
  mutate(
    Performance = factor(Performance, levels = c("exceeds", "meets", "developing", "unsatisfactory"), 
    labels = c("Exceeds Expectations", "Meets Expectations", "Developing", "Unsatisfactory")),
    indicator = case_when(
      str_detect(indicator, "Accountability") ~ "Accountability",
      str_detect(indicator, "Belonging") ~ "Belonging",
      str_detect(indicator, "Objectives") ~ "Objectives",
      TRUE ~ indicator
    ),
    indicator = factor(indicator, levels = c("Accountability", "Belonging", "Objectives")),
    course = case_when(
      str_detect(course, "CCE 102") ~ "CCE 102",
      str_detect(course, "CE 291") ~ "CE 291R",
      str_detect(course, "CE 472") ~ "CE 472",
      TRUE ~ NA_character_
    ) 
  ) |> 
    group_by(course, indicator) |>
    mutate(total = sum(Number)) |>
    mutate(Percent = if_else(total > 0, Number / total * 100, NA_real_))
```

We assess this student outcome in the following courses with the assessments listed in @tbl-so5-instruments.
Complete instrument prompts are available in the program assessment folder.

```{r}
#| label: tbl-so5-instruments
#| tbl-cap: Assessment Instruments for Student Outcome 5
so5_instruments <- so5_clean |> 
  group_by(course, indicator, description) |>
  slice(1) |>
  mutate(level = ifelse(course == "CE 472", "Summative", "Formative"))  |>
  arrange(indicator, level, course) |>
  ungroup() |>
  select(indicator, level, course, description) |> 
  setNames(c("indicator", "Assessment Level", "Course", "Instrument")) |>
  format_table_for_output()

so5_group_rows <- so5_instruments |>
  count(indicator) |>
  mutate(row = cumsum(lag(n, default = 0) + 1)) |>
  pull(row)

tt(so5_instruments)  |>
  group_tt(i = indicator) |>
  subset(select = -indicator)  |> 
  style_tt(align = "l") |>
  style_tt(
    i = so5_group_rows,
    color = "white",
    background = "gray",
    bold = TRUE)

```

### Data and Evaluation

We evaluated this outcome most recently in Winter 2026. Current data include
formative assessment in CCE 102 and summative assessment in CE 472. @fig-s05
shows the share of students at each performance level for each indicator for
each course, alongside the threshold of 75%. Students in CCE 102 met the
threshold for all three indicators. Students in CE 472 were below the threshold
for all three indicators, with 69 percent of students meeting or exceeding
expectations for each indicator.

```{r}
#| label: fig-s05
#| fig-cap: Fall 2025 and Winter 2026 Assessment of Student Outcome 5
#| fig-width: 7
ggplot(so5_clean, aes(y = Percent, x = course, fill = fct_rev(Performance))) + 
  geom_col() + 
  labs(x = "Course", y = "Percent of students") +  facet_grid(~ indicator) + 
  geom_hline(yintercept = 75, linetype = "dashed", color = "red") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual("Performance", values = pal)
```


## Experiments and Data {#sec-so6}

Students will demonstrate:

> An ability to develop and conduct appropriate experimentation, analyze and
> interpret data, and use engineering judgment to draw conclusions.

### Performance Indicators {#sec-so6-performance}

We evaluate this outcome based on four performance indicators. Students should:

1. Identify conditions: Identify what conditions need to be controlled or varied in an experiment in order to obtain data needed to test a hypothesis. 
2. Select methods: Select methods, materials, and/or equipment that will generate the desired data. 
3. Conduct: Conduct the experiment and keep appropriate records. 
4. Analyze and conclude: Analyze data using scientific techniques and draw supported conclusions. 

We assess these performance indicators with the rubric in @tbl-so6-rubric.

| Indicator                | Unsatisfactory                                                                                                                     | Developing                                                                                                                                     | Meets Expectations                                                                                                                                   | Exceeds expectations                                                                                                                                                                                         |
|:-------------------------|:-----------------------------------------------------------------------------------------------------------------------------------|:-----------------------------------------------------------------------------------------------------------------------------------------------|:-----------------------------------------------------------------------------------------------------------------------------------------------------|:-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------|
| 1. Identify conditions | Does not distinguish controls and variables; attempts multiple variables in one experiment                                         | Identifies some controls and variables                                                                                                         | Correctly identifies all controls and variables                                                                                                      | Correctly identifies all controls and variables as well as possible confounding conditions                                                                                                                   |
| 2. Select methods | Does not match methods to experiment                                                                                               | Selects some adequate methods, materials, or equipment                                                                                         | Selects adequate methods, materials, and equipment; can justify choice                                                                               | Selects adequate methods, materials, and equipment; can justify choice; may use standards                                                                                                                    |
| 3. Conduct | Careless setup or execution of experiment; little or no recordkeeping; inattentive to experiment                                   | Some problems in setup or execution; keeps some records                                                                                        | Carefully sets up and executes experiment; records include photos, notes, observations, and potential errors                                         | Carefully sets up and executes experiment; records include photos, notes, observations, and potential errors or limitations                                                                                  |
| 4. Analyze and conclude | Little or no analysis of data; incorrect analysis of data; no conclusion; concludes without data; concludes contradictory to data  | Shows some analysis by figures, equations, statistics, etc., though perhaps with mistakes; may conclude without data or contradictory to data  | Demonstrates thorough analysis by figures, equations, statistics, etc.; concludes based on data and in accordance with data; recognizes assumptions  | Demonstrates thorough analysis by figures, equations, statistics, etc.; concludes based on data and in accordance with data; recognizes assumptions; acknowledges civil engineering applications of results  |
: Rubric to assess student outcome 6: Experiments and Data {#tbl-so6-rubric}

### Assessment Instruments {#sec-so6-assessment}

We assess this student outcome at three levels:

- CCE 270: Computer Methods (introductory)
- CE 232: Fluid Mechanics (reinforcing)
- CE Breadth classes (evaluation): CCE 306 Civil Engineering Materials, CE 341 Soil Mechanics, and CE 351 Environmental Engineering

All CE students must take one of the three breadth classes where we assess this outcome, 
though many will take more than one.

```{r so6-clean, include = FALSE}
so6 <- read_assessment_data("data/Student Outcome 6_ Experiments and Data.csv")
so6_clean <- so6 |>
  mutate(
    course = str_squish(course),
    assessment = na_if(str_squish(assessment), ""),
    level = case_when(
      course %in% c("CCE 270", "CE 232") ~ "Formative",
      course %in% c("CCE 306", "CE 341", "CE 351") ~ "Summative",
      TRUE ~ NA_character_
    ),
    course = factor(course, levels = c("CCE 270", "CE 232", "CCE 306", "CE 341", "CE 351")),
    indicator = case_when(
      indicator == "identify" ~ "Identify Conditions",
      indicator == "select" ~ "Select Methods",
      indicator == "conduct" ~ "Conduct",
      indicator == "analyze" ~ "Analyze and Conclude",
      TRUE ~ indicator
    ),
    indicator = factor(
      indicator,
      levels = c("Identify Conditions", "Select Methods", "Conduct", "Analyze and Conclude")
    )
  ) |>
  pivot_longer(exceeds:unsatisfactory, values_to = "Percent", names_to = "Performance") |>
  filter(!is.na(Percent)) |>
  mutate(
    Performance = factor(Performance, 
      levels = c("exceeds", "meets", "developing", "unsatisfactory"), 
      labels = c("Exceeds Expectations", "Meets Expectations", "Developing", "Unsatisfactory")
    )
  )
```

We assess this student outcome in the following courses with the assessments listed in @tbl-so6-instruments.

```{r}
#| label: tbl-so6-instruments
#| tbl-cap: Assessment Instruments for Student Outcome 6
so6_instruments <- so6_clean |>
  group_by(course, indicator, assessment, level) |>
  slice(1) |>
  arrange(indicator, level, course) |>
  ungroup() |>
  mutate(Instrument = coalesce(assessment, "Assessment not specified")) |>
  select(indicator, level, course, Instrument) |>
  setNames(c("indicator", "Assessment Level", "Course", "Instrument")) |>
  format_table_for_output()

so6_group_rows <- so6_instruments |>
  count(indicator) |>
  mutate(row = cumsum(lag(n, default = 0) + 1)) |>
  pull(row)

tt(so6_instruments) |>
  group_tt(i = indicator) |>
  subset(select = -indicator) |>
  style_tt(align = "l") |>
  style_tt(
    i = so6_group_rows,
    color = "white",
    background = "gray",
    bold = TRUE)

```

### Data and Evaluation

The current data include Winter 2025 assessments in CE 232 and the breadth
courses and Winter 2026 assessment in CCE 270. @fig-s06 shows the share of
students assessed at each performance level for each indicator in the relevant
courses. The CE 232 formative assessment meets the threshold for all four
indicators. The CCE 270 formative assessments are just below the threshold for
Select Methods and Conduct. Summative breadth-course evidence is uneven: CCE 306
and CE 341 meet the threshold for Conduct, CE 341 meets the threshold for
Identify Conditions, and the remaining summative assessments are below the
threshold.

```{r}
#| label: fig-s06
#| fig-cap: Winter 2025 and Winter 2026 Assessment of Student Outcome 6
#| fig-width: 7
ggplot(so6_clean, aes(y = Percent, x = course, fill = fct_rev(Performance))) +
    geom_col() +
    labs(x = "Course", y = "Percent of students") +
    facet_grid(~indicator) +
    geom_hline(yintercept = 75, linetype = "dashed", color = "red") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
scale_fill_manual("Performance", values = pal)
```

The committee recommends calibration among the breadth-course laboratory
instructors, with particular attention to Select Methods and Analyze and
Conclude, so that expectations and scoring are consistent across the courses
used for summative assessment.

## Acquire Knowledge {#sec-so7}

Students will demonstrate:

> An ability to acquire and apply new knowledge as needed, using appropriate
> learning strategies.

### Performance Indicators {#sec-so7-performance}

We evaluate this outcome based on three performance indicators. Students should:

1. Effectively seek out appropriate resources to acquire knowledge and understanding of unfamiliar concepts.
2. Critically evaluate sources for reputability and reliability.
3. Appropriately apply newly acquired knowledge and skills to complete an analysis.

We assess these performance indicators with the rubric in @tbl-so7-rubric.

| Indicator | Unsatisfactory | Developing | Meets Expectations | Exceeds Expectations |
|:---|:---|:---|:---|:---|
| 1. Seek resources | Relies exclusively on provided materials; shows little or no initiative in finding new resources. | Consults additional resources when prompted and demonstrates partial understanding of how the resources contribute to learning. | Independently identifies and uses credible, relevant resources that contribute meaningfully to understanding. | Seeks diverse, high-quality resources and synthesizes information from multiple sources to deepen understanding. |
| 2. Evaluate sources | Resources used, if any, are irrelevant, unreliable, or misinterpreted. | Uses resources with limited evaluation of quality or relevance. | Provides evidence of evaluating resource quality and applicability. | Critically analyzes the strengths and limitations of sources. |
| 3. Apply knowledge | Fails to apply newly acquired techniques, or applies them incorrectly, demonstrating misunderstanding or lack of effort to learn. | Attempts to use new techniques but with significant errors or incomplete understanding; requires guidance to complete the analysis. | Correctly applies newly learned techniques and practices to perform an appropriate analysis with minimal errors. | Applies new techniques accurately and efficiently, extending them beyond the basic requirements. |
: Student Outcome 7 Assessment Rubric {#tbl-so7-rubric}

### Assessment Instruments {#sec-so7-assessment}

```{r so7-clean, include = FALSE}
so7 <- read_assessment_data("data/Student Outcome 7_ Learning and Service(Sheet1).csv")

so7_clean <- so7 |>
  pivot_longer(exceeds:unsatisfactory, names_to = "Performance", values_to = "Number") |>
  filter(!is.na(Number)) |>
  mutate(
    Performance = factor(
      Performance,
      levels = c("exceeds", "meets", "developing", "unsatisfactory"),
      labels = c("Exceeds Expectations", "Meets Expectations", "Developing", "Unsatisfactory")
    ),
    indicator = case_when(
      str_detect(str_to_lower(indicator), "seeks out") ~ "Seek Resources",
      str_detect(str_to_lower(indicator), "evaluates sources") ~ "Evaluate Sources",
      str_detect(str_to_lower(indicator), "applies newly acquired") ~ "Apply Knowledge",
      TRUE ~ indicator
    ),
    indicator = factor(indicator, levels = c("Seek Resources", "Evaluate Sources", "Apply Knowledge")),
    course = case_when(
      str_detect(course, "CCE 113") ~ "CCE 113",
      str_detect(course, "CCE 114") ~ "CCE 114",
      str_detect(course, "CE 471") ~ "CE 471",
      TRUE ~ NA_character_
    ),
    course = factor(course, levels = c("CCE 113", "CCE 114", "CE 471"))
  ) |>
  group_by(course, indicator) |>
  mutate(total = sum(Number)) |>
  mutate(Percent = if_else(total > 0, Number / total * 100, NA_real_))
```

We assess this student outcome in the following courses with the assessments listed in @tbl-so7-instruments.
Complete instrument prompts are available in the program assessment folder.

```{r}
#| label: tbl-so7-instruments
#| tbl-cap: Assessment Instruments for Student Outcome 7
so7_instruments <- so7_clean |>
  group_by(course, indicator, description) |>
  slice(1) |>
  mutate(level = ifelse(course == "CE 471", "Summative", "Formative")) |>
  arrange(indicator, level, course) |>
  ungroup() |>
  select(indicator, level, course, description) |>
  setNames(c("indicator", "Assessment Level", "Course", "Instrument")) |>
  format_table_for_output()

so7_group_rows <- so7_instruments |>
  count(indicator) |>
  mutate(row = cumsum(lag(n, default = 0) + 1)) |>
  pull(row)

tt(so7_instruments) |>
  group_tt(i = indicator) |>
  subset(select = -indicator) |>
  style_tt(align = "l") |>
  style_tt(
    i = so7_group_rows,
    color = "white",
    background = "gray",
    bold = TRUE)

```

### Data and Evaluation

We evaluated this outcome most recently in Winter 2026. Current data include
formative assessment in CCE 113 and CCE 114 and summative assessment in CE 471.
@fig-s07 shows the share of students at each performance level for each
indicator for each course, alongside the threshold of 75%. The CE 471 summative
assessment meets the threshold for Seek Resources and Evaluate Sources. The
current data file does not include a nonzero summative assessment for Apply
Knowledge; formative assessment of Apply Knowledge in CCE 113 and CCE 114 is
slightly below the threshold.

```{r}
#| label: fig-s07
#| fig-cap: Fall 2025 and Winter 2026 Assessment of Student Outcome 7
#| fig-width: 7
ggplot(so7_clean, aes(y = Percent, x = course, fill = fct_rev(Performance))) +
  geom_col() +
  labs(x = "Course", y = "Percent of students") +
  facet_grid(~ indicator) +
  geom_hline(yintercept = 75, linetype = "dashed", color = "red") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual("Performance", values = pal)
```

The committee recommends collecting complete summative evidence for all three
Acquire Knowledge indicators in the next assessment cycle, with particular
attention to Apply Knowledge.
