# Student Outcomes {#sec-so}

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(sharepointr)
library(kableExtra)
get_file <- function(url, outcome, indicator, course) {
  folder <- file.path("assessments", outcome, indicator) 
  dir.create(folder, recursive = TRUE, showWarnings = FALSE)
  download.file(url, file.path(folder, paste0(course, ".pdf")))
}

library(wesanderson)
pal <- wes_palette("AsteroidCity1", 4)[c(3, 2, 1, 4)]
theme_set(theme_minimal())
```

The program assesses each of its seven student outcomes using performance
indicators and associated rubrics.


::: {.panel-tabset}
## SO 1 {#sec-so1}



## SO 2 {#sec-so2}


## SO 3 {#sec-so3}


## SO 4 {#sec-so4}
An ability to recognize ethical and professional responsibilities in engineering situations and make informed judgments, which must consider the impact of engineering solutions in global, economic, environmental, and societal contexts.

### Performance Indicators {.unnumbered}

We evaluate this outcome based on three performance indicators. Students should:

1.	Recognize personal and professional responsibilities towards others including employees, clients, the wider profession, and the public.
2.	Recognize moral and ethical dilemmas in engineering situations and make judgments to identify appropriate resolutions.
3.	Consider the impacts of an engineering solution in global, economic, environmental, and societal contexts.

### Rubric {.unnumbered}

| Indicator                  | Unsatisfactory                                                                | Developing                                                                           | Meets Expectations                                                             | Exceeds expectations                                                                                                                |
|----------------------------|-------------------------------------------------------------------------------|--------------------------------------------------------------------------------------|--------------------------------------------------------------------------------|-------------------------------------------------------------------------------------------------------------------------------------|
| Recognize responsibilities | Does not recognize or identify professional responsibilities.                 | Recognizes some professional responsibilities, but misses at least one.              | Recognizes professional responsibilities towards all key stakeholders          | Demonstrates interest in and  motivation to serve society through the profession.                                                   |
| Make judgments             | Does not identify ethical dilemmas, and or develops inappropriate resolutions | Misidentifies ethical dilemmas, and or develops inadequate resolutions.              |  Properly identifies ethical dilemmas and identifies appropriate resolutions.  | Actively considers ethical challenges that may arise in professional contexts and seeks to mitigate  them.                          |
| Consider impacts           | Fails to identify impacts in any contexts.                                    | Identifies impacts in some contexts, but fails to identify likely impacts in others. | Identifies likely impacts in a variety  of contexts.                           | Actively seeks to consider the ethical impact of engineering situations in global, economic, environmental, and societal  contexts. |

### Assessment {.unnumbered}

We assess this learning outcome in three courses: 

  - CCE 102: Sustainable Infrastructure (Introductory)
  - CE 291 R: Civil Engineering Seminar (Reinforcing)
  - CE 472: Civil Engineering Capstone (Evaluation)

We evaluated this outcome most recently in Winter 2025. All graduating students take CE 472 in Winter semester. Students take CE 291 R each semester from their sophomore year onward. CCE 102 may be taken in Fall or Winter semesters, so this is approximately a 50% sample of the students in the program. CCE 102 also includes students who are in the CFM and the FM programs.

```{r so4-clean, include = FALSE}
so4 <- read_csv("data/Student Outcome 4_ Ethics, Faith and Sustainability(Sheet1).csv") |> 
  set_names(c("id", "start_time", "completion_time", "email", "name", "course", "description", "prompt", 
    "indicator", "exceeds", "meets", "developing", "unsatisfactory", "exceeds2", "meets2", "developing2", "unsatisfactory2",
    "exceeds3", "meets3", "developing3", "unsatisfactory3")) |> 
  select(id:unsatisfactory) |> 
  pivot_longer(exceeds:unsatisfactory, names_to = "Performance", values_to = "Number") |>
  mutate(
    Performance = factor(Performance, levels = c("exceeds", "meets", "developing", "unsatisfactory"), 
    labels = c("Exceeds Expectations", "Meets Expectations", "Developing", "Unsatisfactory")),
    indicator = factor(indicator, labels = c("Consider Impacts", "Make Judgments", "Recognize Responsibilities")),
    course = case_when(
      str_detect(course, "CCE 102") ~ "CCE 102",
      str_detect(course, "CE 291") ~ "CE 291",
      str_detect(course, "CE 472") ~ "CE 472",
      TRUE ~ NA_character_
    ) 
  ) |> 
    group_by(course, indicator) |>
    mutate(total = sum(Number)) |>
    mutate(Percent = Number / total * 100)
```

@fig-s04 shows the share of students at each performance level for each indicator for each course, 
alongside the threshold of 75%. By taking CE 472, 75% of students meet the expectations for each indicator.

```{r}
#| label: fig-s04
#| fig-cap: Winter 2025 Assessment of Student Outcome 4
#| fig-width: 7
ggplot(so4, aes(y = Percent, x = course, fill = fct_rev(Performance))) + 
  geom_col() + 
  labs(x = "Percent of students", y = "Performance") +
  facet_grid(~ indicator) + 
  geom_hline(yintercept = 75, linetype = "dashed", color = "red") +
  scale_fill_manual("Performance", values = pal)
```

```{r get-files, eval = FALSE}
# download the files in the pdf document field
for(i in 1:nrow(so4)) {
  get_file(
    url = so4$`Upload a PDF document that contains the assessment prompt or a complete description of the assignment.`[i],
    outcome = "SO4",
    indicator = so4$`Which performance indicator does this assessment address?`[i],
    course = so4$`Which course is this assessment given in?`[i]
  )
}

```



## SO 5 {#sec-so5}


## SO 6 {#sec-so6}


## SO 7 {#sec-so7}

:::



## Assessment Schedule

The program assesses each student outcome on a rolling three-year cycle.

