# Exit Surveys and Interviews {#sec-exit}
```{r setup, include=FALSE}
library(tidyverse)
library(tinytable)
library(haven)
library(alluvial)
```

## Exit Surveys {#sec-survey}
Each December and April, the program sends a survey to all of our graduating
students. The purpose of this survey is to collect data on the following items:

  1. Self-assessment of student outcomes
  2. Prospective employment or graduate school plans
  3. Belonging and extracurricular involvement
  4. Program Evaluation

```{r load-survey-data, include = FALSE}
survey_data <- read_sav("data/survey/2026+CE+Undergraduate+Exit+Survey_June+4%2C+2026_15.14.sav") |>
  mutate(year = lubridate::year(StartDate)) 

```

In Winter 2026, the committee streamlined and overhauled the survey to give more targeted answers
and avoid duplication with surveys administered by BYU Alumni. `r nrow(survey_data)` students
responded to the 2026 survey.

### Self-assessment of student outcomes

The survey asks students to self-assess their preparedness on the student
outcomes and related preparation categories described in @sec-so along a
five-point scale from "Not at all prepared" to "Very well prepared." Two outcomes
are divded into separate categories: SO3 Communication is divided into written and
oral communication, and SO2 Design is divided into design solutions and design impacts 
(as in the "consider factors" indicator).


```{r}
#| label: fig-exit-prepared-lo
#| fig-cap: Student self-assessment response proportions by preparation category
#| fig-width: 8
prepared_lo_categories <- c(
  prepared_lo_1 = "Complex problems",
  prepared_lo_2 = "Design solutions",
  prepared_lo_3 = "Design impacts",
  prepared_lo_4 = "Written communication",
  prepared_lo_5 = "Oral communication",
  prepared_lo_6 = "Ethics",
  prepared_lo_7 = "Teamwork",
  prepared_lo_8 = "Data analysis",
  prepared_lo_9 = "Learn independently",
  prepared_lo_10 = "Overall preparation"
)

preparedness_levels <- c(
  "1 - Not at all prepared",
  "2 - Slightly prepared",
  "3 - Moderately prepared",
  "4 - Well prepared",
  "5 - Very well prepared"
)

prepared_lo_counts <- survey_data |>
  select(all_of(names(prepared_lo_categories))) |>
  mutate(across(everything(), as.numeric)) |>
  pivot_longer(everything(), names_to = "category", values_to = "preparedness") |>
  filter(!is.na(preparedness)) |>
  mutate(
    category = factor(
      category,
      levels = names(prepared_lo_categories),
      labels = prepared_lo_categories
    ),
    preparedness = factor(
      preparedness,
      levels = 1:5,
      labels = preparedness_levels
    )
  ) |>
  count(category, preparedness) |>
  complete(category, preparedness, fill = list(n = 0)) |>
  group_by(category) |>
  mutate(proportion = n / sum(n)) |>
  ungroup()

ggplot(prepared_lo_counts, aes(x = category, y = proportion, fill = preparedness)) +
  geom_col(width = 0.7) +
  geom_hline(yintercept = 0.75, linetype = "dashed", color = "red") +
  labs(x = "Preparation category", y = "Proportion of responses", fill = "Preparedness") +
  scale_y_continuous(
    labels = scales::percent_format(),
    breaks = seq(0, 1, 0.25),
    limits = c(0, 1)
  ) +
  scale_fill_brewer(type = "seq", palette = "Blues", direction = 1) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

In only one category (Learn independently) do at least 75% of students rate their
preparation as "Well prepared" or higher. The categories with the most potential 
for growth in this area are Compmlex Problems and Ethics.

### Prospective employment or graduate school plans

The survey asks students to report their immediate plans after graduation and,
where applicable, their future graduate school plans. @fig-exit-plans-alluvial 
shows the distribution students across these levels. More than half of the students
suggest that they will pursue some form of graduate education in their future.

```{r}
#| label: fig-exit-plans-alluvial
#| fig-cap: Prospective employment and graduate school plans
#| fig-width: 9
#| fig-height: 6
graduate_degree_labels <- c(
  graduate_which_1 = "MS/MEng CE",
  graduate_which_2 = "PhD CE",
  graduate_which_3 = "MBA",
  graduate_which_4 = "Related field",
  graduate_which_5 = "Different field",
  graduate_which_6 = "Not sure"
)

graduate_degree_stage <- survey_data |>
  select(all_of(names(graduate_degree_labels))) |>
  mutate(across(everything(), ~ !is.na(.x))) |>
  pmap_chr(function(...) {
    selected <- names(graduate_degree_labels)[as.logical(c(...))]

    if (length(selected) == 0) {
      return("No degree")
    }

    if (length(selected) > 1) {
      return("Multiple types")
    }

    unname(graduate_degree_labels[selected])
  })

prospective_flows <- survey_data |>
  mutate(
    plan_response = as.character(as_factor(plans)),
    future_response = as.character(as_factor(graduate_future)),
    where_response = as.character(as_factor(graduate_where)),
    plan = case_when(
      plan_response == "Work in civil engineering" ~ "Work in CE",
      plan_response == "Graduate school, civil engineering" ~ "Grad school: CE",
      plan_response == "Graduate school, other field" ~ "Grad school: other",
      plan_response == "Other (please specify)" ~ "Other",
      TRUE ~ NA_character_
    ),
    graduate_future = case_when(
      str_detect(plan, "^Grad school") ~ "Immediate grad school",
      !is.na(future_response) ~ future_response,
      TRUE ~ "No response"
    ),
    graduate_degree = graduate_degree_stage,
    graduate_where = case_when(
      str_detect(where_response, "Here at BYU") ~ "BYU",
      str_detect(where_response, "At another university") ~ "Another university",
      str_detect(where_response, "Not sure") ~ "Not sure",
      TRUE ~ "No location"
    ),
    plan = factor(plan, levels = c("Work in CE", "Grad school: CE", "Grad school: other", "Other")),
    graduate_future = factor(
      graduate_future,
      levels = c("Immediate grad school", "Yes", "No", "Not sure yet", "No response")
    ),
    graduate_degree = factor(
      graduate_degree,
      levels = c(
        "MS/MEng CE", "PhD CE", "MBA", "Related field", "Different field",
        "Not sure", "Multiple types", "No degree"
      )
    ),
    graduate_where = factor(
      graduate_where,
      levels = c("BYU", "Another university", "Not sure", "No location")
    )
  ) |>
  filter(!is.na(plan)) |>
  count(plan, graduate_future, graduate_degree, graduate_where, name = "n")

plan_colors <- c(
  "Work in CE" = "#2C7FB8",
  "Grad school: CE" = "#31A354",
  "Grad school: other" = "#756BB1",
  "Other" = "#E6550D"
)

old_par <- par(no.readonly = TRUE)
par(mar = c(2, 2, 3, 2), xpd = NA)
alluvial::alluvial(
  as.data.frame(prospective_flows[, c("plan", "graduate_future", "graduate_degree", "graduate_where")]),
  freq = prospective_flows$n,
  col = unname(plan_colors[as.character(prospective_flows$plan)]),
  border = "white",
  alpha = 0.75,
  blocks = FALSE,
  axis_labels = c("Immediate plan", "Future graduate study", "Degree type", "Graduate location"),
  cex = 0.65,
  cex.axis = 0.75
)
par(old_par)
```




### Belonging and extracurricular involvement

The survey asks students to evaluate several elements of belonging and to report
the number of meaningful learning experiences they completed during their time
at BYU. 

@fig-exit-belonging shows the distribution of students' responses to the belonging
questions. Student generally report strong levels of agreement with the relevant
statements. There are three statements however, where less than 75% percent of
students report "Agree" or "Strongly agree":

  2. I feel a connection to the Civil & Construction Engineering department.
  2. I feel a connection to the College of Engineering.
  2. The Civil Engineering program provided meaningful opportunities for involvement outside the classroom.


```{r}
#| label: fig-exit-belonging
#| fig-cap: Student agreement with belonging statements
#| fig-width: 8
#| fig-height: 6
belonging_categories <- c(
  belonging__1 = "Belonging in CE",
  belonging__2 = "Faculty support",
  belonging__3 = "Student support",
  belonging__4 = "Civil engineer identity",
  belonging__5 = "CE program connection",
  belonging__6 = "Department connection",
  belonging__7 = "College connection",
  belonging__8 = "BYU connection",
  belonging__9 = "Involvement opportunities",
  belonging__10 = "Social connection"
)

agreement_levels <- c(
  "Strongly disagree",
  "Disagree",
  "Neutral",
  "Agree",
  "Strongly agree"
)

belonging_counts <- survey_data |>
  select(all_of(names(belonging_categories))) |>
  mutate(across(everything(), as.numeric)) |>
  pivot_longer(everything(), names_to = "element", values_to = "agreement") |>
  filter(!is.na(agreement)) |>
  mutate(
    element = factor(
      element,
      levels = rev(names(belonging_categories)),
      labels = rev(belonging_categories)
    ),
    agreement = factor(
      agreement,
      levels = 1:5,
      labels = agreement_levels
    )
  ) |>
  count(element, agreement) |>
  complete(element, agreement, fill = list(n = 0))

ggplot(belonging_counts, aes(x = element, y = n, fill = agreement)) +
  geom_col(position = "fill", width = 0.75) +
  geom_hline(yintercept = 0.75, linetype = "dashed", color = "red") +
  coord_flip() +
  labs(x = NULL, y = "Proportion of responses", fill = "Agreement") +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_brewer(type = "div", palette = "Blues") +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 1))
```


The university has a goal for students to complete at least two learning experiences 
during their time at BYU. @fig-exit-learning-experiences shows the distribution of
the number of learning experiences reported by students. Completing the capstone is a requirement
of graduation, though some students did not mark having experienced this. 


```{r}
#| label: fig-exit-learning-experiences
#| fig-cap: Distribution of learning experiences reported by type
#| fig-width: 8
#| fig-height: 6
learning_experience_categories <- c(
  learning_experiences_1_1 = "Culminating learning experience",
  learning_experiences_2_1 = "Field study/field work",
  learning_experiences_3_1 = "Internship",
  learning_experiences_4_1 = "Mentored research",
  learning_experiences_5_1 = "Study abroad",
  learning_experiences_6_1 = "Campus engagement"
)

learning_experience_counts <- survey_data |>
  select(all_of(names(learning_experience_categories))) |>
  mutate(
    across(everything(), ~ readr::parse_number(as.character(.x))),
    across(everything(), ~ ifelse(is.na(.x), 0, .x))
  )  |>
  mutate(total_experiences = rowSums(across(everything()), na.rm = TRUE)) |>
  pivot_longer(everything(), names_to = "experience_type", values_to = "experiences") |>
  mutate(
    experience_type = factor(
      experience_type,
      levels = c(names(learning_experience_categories), "total_experiences"),
      labels = c(learning_experience_categories, "Total Experiences")
    )
  ) |> 
  count(experience_type, experiences, name = "respondents") |>
  mutate(
    experiences = ifelse(experiences > 7, 7, experiences)
  )

ggplot(learning_experience_counts, aes(x = experiences, y = respondents)) +
  geom_col(fill = "#2C7FB8", width = 0.8) +
  facet_wrap(~ experience_type, ncol = 2) +
  labs(x = "Experiences reported", y = "Number of respondents") +
  scale_x_continuous(breaks = scales::breaks_width(1)) +
  scale_y_continuous(breaks = scales::breaks_width(1))
```

### Program Evaluation



## Exit Interviews {#sec-interview}

Each April, the undergraduate committee samples graduating students and
interviews them in personal conversation. The purpose of this interview is to
qualitatively assess student-observed program strengths and weakness.

The CE undergraduate coordinator assigns three to four graduating seniors to
each member of the committee. The committee members arrange to speak privately
with each student for fifteen to twenty minutes. The student should be invited
to direct the discussion, but possible questions include: 

  - What are the most positive learning experiences you have had?
  - Have you had any less successful learning experiences?

Each committee member makes notes of the student's responses. The assessmentcoordinator
consolidates these notes into a coherent statment and posts it in the space below.

::: {.panel-tabset}
### Class of 2026

The civil engineering program demonstrates several clear strengths, particularly in the quality of instruction in upper-division courses and the accessibility of faculty. Students consistently report positive experiences with professors who are engaged, approachable, and effective at connecting course material to professional practice. This is especially evident in areas such as structures, geotechnical engineering, and transportation, where well-organized courses and strong instruction enhance learning outcomes. The program also supports successful career placement, with graduates securing relevant employment, internships, and graduate school opportunities. High-impact experiences such as capstone projects, undergraduate research, and study abroad further contribute to student development. Additionally, the program fosters a strong sense of community among students, particularly in later years, where stable cohorts and peer interaction enhance the overall educational experience.

Despite these strengths, the program exhibits several systemic weaknesses related to coherence, structure, and student engagement. The introductory experience is inconsistent and, in some cases, ineffective at orienting students to the discipline or fostering a sense of belonging, particularly for transfer students. There are also notable gaps between foundational coursework and upper-division applications, including insufficient alignment between mathematics preparation and engineering problem-solving. Students report variability in course organization and grading practices, as well as occasional overemphasis on tools (e.g., programming) at the expense of conceptual understanding. The seminar program is widely viewed as inefficient in its current form, with diminishing value in upper-division years and a need for greater emphasis on career preparation and interactive engagement. Similarly, the capstone experience lacks clarity and alignment with prior coursework, and is perceived by some students as placing undue focus on low-value tasks. Finally, student participation in co-curricular activities is inconsistent, suggesting that current approaches to engagement may be insufficiently effective.

### Class of 2025
Students described the civil engineering program as offering strong overall
preparation, with a solid breadth of coursework and particularly effective
introductory classes that sparked interest in the major. Many felt well prepared
for the FE exam and appreciated practical, real-world applications in courses
like CE 471, as well as hands-on experiences through competition teams,
research, study abroad, and TA roles. Faculty were consistently viewed as
enthusiastic, accessible, and invested in student success, especially in smaller
upper-division classes where interaction was more direct. The program fostered a
strong sense of belonging through clubs, study groups, and shared spaces,
supported by active (though sometimes inconsistent) faculty involvement.
Students valued flexible curriculum elements like the junior core options, and
many appreciated opportunities to integrate faith meaningfully into their
education. Physical spaces such as labs and common areas also contributed
positively to collaboration and community.

At the same time, students identified several persistent challenges. Seminar was
widely seen as repetitive and inefficient, with too many guest speakers and not
enough practical skill development, and its scheduling sometimes conflicted with
internships. There were concerns about uneven course quality, disruptions from
instructor turnover, limited TA support in upper-division courses, and heavy
workloads that restricted extracurricular involvement. Technical skills like
CAD, Revit, and Excel were not consistently reinforced, leading to skill loss
over time. Writing requirements—particularly ENGL 316—were viewed as overly
burdensome and less relevant than discipline-specific alternatives like CE 471,
with some students suggesting earlier placement in the curriculum. Capstone
grading was sometimes perceived as overly nitpicky, and some assignments felt
like busywork. Students also noted scheduling limitations (e.g., course
conflicts, weaker winter offerings), a lack of advanced or graduate-level
electives, and a need for clearer guidance on electives and career pathways.
Additional suggestions included improving inclusivity and representation
(especially for women), expanding interdisciplinary and coding opportunities,
increasing faculty engagement in student organizations, and addressing
administrative friction points such as curriculum changes and the graduate
application process.
:::
