# Fundamentals of Engineering {#sec-fe}
```{r setup-fe, include=FALSE}
library(tidyverse)
library(readxl)

knitr::opts_chunk$set(echo = FALSE)
semester_to_date <- function(semester) {
  year <- as.numeric(stringr::str_extract(semester, "[0-9]+"))
  type <- stringr::str_extract(semester, "[a-zA-Z]+")
  
  date_str <- dplyr::case_when(
    type == "Winter" ~ stringr::str_c(year, "-06-01"),
    type == "Fall"   ~ stringr::str_c(year, "-12-01"),
    TRUE             ~ NA_character_
  )
  
  lubridate::ymd(date_str)
}

theme_set(theme_bw())
pal <- c("#FF6127", "#002E5D")
```

The Fundamentals of Engineering (FE) examination provides important external
evidence of our program's effectiveness in training competent civil engineering
professionals. University policies prohibit us from requiring students to take
the exam before graduation, but we strongly recommend that our students do so
and offer an FE preparation course to help them.

```{r fe}
fe <- readxl::read_xlsx("data/fe.xlsx", sheet = "raw_total") |>
    mutate( period = semester_to_date(period))
```

The share of students passing the FE exam is shown in @fig-fe-exam. BYU students
pass the exam at a rate considerably higher than other ABET-accredited civil engineering programs. This data
is for students currently enrolled in the program, or graduating in the 12
months prior to taking the exam. We do not include alumni who take the FE more
than one year after graduation in this assessment.

```{r}
#| label: fig-fe-exam
#| fig-cap: FE Exam pass rate by semester, BYU Civil and ABET comparator group.
fe |>
    filter(category == "Pct Passing") |>
    pivot_longer(cols = c(`BYU-Civil`, `ABET-Civil`), names_to = "group", values_to = "pct_passing") |>
    ggplot(aes(x = period, y = pct_passing, color = group)) +
    geom_line() + scale_y_continuous(labels = scales::percent)  +
    xlab("Period of Examination [six months ending]") + ylab("Percent Passing") + scale_color_manual("Group", values = pal)
```


We also use the FE exam to assess discipline and topic subscores. @fig-fe-subscores provides an
interactive graphic where you can choose an FE exam category and compare BYU civil engineering students
and recent alumni against an "ABET comparator" group of similar institutions. BYU students generally exceed 
the ABET comparator group mean in all categories for almost all periods, but fall within a single standard
deviation of that mean. There are a few exceptions to this performance, with BYU students modestly below the ABET
comparator mean in some categories for some periods. These include the "Ethics and Professional Practice" and the 
"Construction Engineering" categories.

::: {#fig-fe-subscores}
```{r fe-shiny}
if(knitr::is_html_output()){
  knitr::include_app("https://019d4550-8785-50dd-0b75-24fc8358dff6.share.connect.posit.cloud/", height = "600px")
} else {

}
```

 FE Exam category subscores from Fall 2022 onwards. Interactive application on [ShinyApps](https://019d4550-8785-50dd-0b75-24fc8358dff6.share.connect.posit.cloud/).
::: 

::: {.callout-note title="Assessment Practices" collapse="true"}
NCEES makes aggregated FE scores available to colleges for assessment purposes. The
College of Engineering at BYU has a login name to this system, which is shared with
the department academic advisor. 

Each August and each February, the assessment coordinator retrieves the most recent sores
from the advisor and adds them to data records in this repository. There are two places where
these data are stored: 

  - The `data/fe.xlsx` spreadsheet contains the data from the advisor's login as well as a worksheet of the 
  general exam pass rates.
  - The `apps/data.csv` files contains the exam subscores in machine-readable format.

To update the report, the assessment coordinator does the following:

  1. Runs the chunks on this page to update the FE pass rate figure.
  2. Generates the shiny app in the `apps` folder with the new data.
  3. Uploads the shiny app to the posit connect cloud and updates the links to the app
    in the report.
  4. Edits the report text to reflect any additional evaluation of the assessment data.
:::